---
title: "Aplied Statistics Project"
author: "José Pinto - 201603713"
date: "5/1/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective 

The objective of this project is to, using linear regression, try to predict the presence<br>
or absence of cardiovascular disease in a particular individual.<br>

With that goal in mind a dataset was obtained containing several features of theoretical<br>
significance.<br>

## Dateset description

The dataset we will be using is the cardiovascular disease dataset hosted in kaggle.<br>
https://www.kaggle.com/sulianova/cardiovascular-disease-dataset/data<br>

The dataset is comprised of about 70000 records of patient data, 11 features and one<br>
target variable.<br>

The features in the dataset are in a few types, whith those being, objective(factual information),<br>
examination(resulting of an exam) or subjective(given by the patient). The type of each<br>
feature will be presented further down.<br>

All values were colected at the moment of medical examination<br>

Note: The source of the dataset could not be verified, as such any results and conclusions<br>
must be carefully analised and its validity can not be fully trusted.<br>

## Dateset import and display

```{r message=FALSE}
library(readr)
cardio_train <- read_delim("C:/Users/zé/Desktop/programação/R/EA/cardio_train.csv", 
                           ";", escape_double = FALSE, trim_ws = TRUE)
knitr::kable(head(cardio_train), caption = "Cardiovascular dataset", floating.environment="sidewaystable")
```

## Feature descriptions

age: The age in days of the person (objective feature), integer value<br>
height: The height in cm of the person(objective feature), integer value<br>
weight: The weight in kg of the person(objective feature), float value<br>
gender: The gender of the individual(objective feature), integer value with categorical<br>
meaning -- 1-women 2-men.<br>
ap-hi: systolic blood pressure(examination feature), integer value<br>
ap-lo: diastolic blood pressure(examination feature), integer value<br>
cholesterol: cholesterol levels of the person(examination feature), integer value with<br>
categorical meaning -- 1-normal 2-above normal 3-well above normal<br>
gluc: glucose levels of the person(examination feature), integer value with categorical<br>
meaning -- 1-normal 2-above normal 3-well above normal<br>
smoke: wether or not the person smokes(subjective feature), boolean value -- 0-does not smoke<br>
1-smokes <br>
alco: wether or not person consumes alcoolic beverages(subjective feature), boolean value<br>
0-no intake 1-intake<br>
active: wether or not person performs physical activity(subjective feature), boolean value<br>
0-no activity 1-activity

### Target variable

cardio: presence or absence of cardiovascular disease, boolean variable, 0-no disease 1-disease<br>

The dataset also has an index row, this will now be removed.<br>

```{r indexRemoval}
cardio_train = cardio_train[2:13]
```

## Means, median, quantiles, min and max

```{r featureSummary}
summary(cardio_train)
```

## Variance

```{r featureVariability}
var(cardio_train$age)
var(cardio_train$gender)
var(cardio_train$height)
var(cardio_train$weight)
var(cardio_train$ap_hi)
var(cardio_train$ap_lo)
var(cardio_train$cholesterol)
var(cardio_train$gluc)
var(cardio_train$smoke)
var(cardio_train$alco)
var(cardio_train$active)
var(cardio_train$cardio)
```

## Features boxplots

```{r featureBoxPlot}
par(mfrow=c(2,2))
boxplot(cardio_train$age,main="age")
boxplot(cardio_train$gender,main="gender")
boxplot(cardio_train$height,main="height")
boxplot(cardio_train$weight,main="weight")
boxplot(cardio_train$ap_hi,main="ap_hi")
boxplot(cardio_train$ap_lo,main="ap_lo")
boxplot(cardio_train$cholesterol,main="cholesterol")
boxplot(cardio_train$gluc,main="gluc")
boxplot(cardio_train$smoke,main="smoke")
boxplot(cardio_train$alco,main="alco")
boxplot(cardio_train$active,main="active")
boxplot(cardio_train$cardio,main="cardio")
```

From the age boxplot we can conclude that most individuals are around 20000 days old ~ 55 years,<br>
with the yougest about 12000 days old ~ 35 years old. The observations are mostly symetric.<br>
From the gender boxplot we can conclude that most of the individuals are female.<br>
From the height boxplot we can conclude that most indivuduals are about 150 to 180 cm tall. We<br>
also have some concerning observations, individuals about 50cm tall, wich is strange given that all<br>
individuals are older that 35 years old. The observations are also skewed to the right.<br>
From the weight boxplot we can conclude that most individuals weigh about 60 kg. Once again we have some<br>
concerning observations of individuals weighing about 20 kg. The observations are also skewed to the right.<br>
From the ap_hi boxplot it is hard to obtain much information other than some severe outliers, with values<br>
reaching the thousands(obviously wrong values).<br>
From the ap_lo boxplot it is hard to obtain much information other than some severe outliers, with values<br>
reaching the thousands(obviously wrong values).<br>
From the cholesterol boxplot we can conclude that most values are of 1, meaning normal levels, with enough<br>
values on the other types for these to not be considered outliers.<br>
From the gluc boxplot we can conclude that most values are of 1, meaning normal levels, however there are<br>
not enough observations on other values, and these are considered outliers.<br>
From the smoke boxplot we can conclude that most individuals are non smokers, with few enough smokers that<br>
these are considered outliers.<br>
From the alco boxplot we can conclude that most individuals are non alcohol consumers, with few enough<br>
consumers that these are considered outliers.<br>
From the active boxplot we can conclude that most individuals are phisicaly active, with few enough<br>
inactive individuals that these are considered outliers.<br>
From the cardio boxplot we can conclude that most individuals don't have cardiovascular diseases, with<br>
enough that have so that these are not considered outliers.<br>

The obviously wrong values spotted in this step will be processed and likely removed at a later time.<br>

## Features histograms

```{r featureHistograms}
par(mfrow=c(2,2))
hist(cardio_train$age,main="age")
gender = table(cardio_train$gender)
barplot(gender, main="Gender Distribution",
  names.arg=c("Male", "Female"))
hist(cardio_train$height,main="height")
hist(cardio_train$weight,main="weight")
hist(cardio_train$ap_hi,main="ap_hi")
hist(cardio_train$ap_lo,main="ap_lo")
cholesterol = table(cardio_train$cholesterol)
barplot(cholesterol, main="Cholesterol Distribution",
  names.arg=c("Normal", "High","Very high"))
gluc = table(cardio_train$gluc)
barplot(gluc, main="Glucose Distribution",
  names.arg=c("Normal", "High","Very high"))
smoke = table(cardio_train$smoke)
barplot(smoke, main="Smoke Distribution",
  names.arg=c("Non smoker", "Smoker"))
alco = table(cardio_train$alco)
barplot(alco, main="Alco Distribution",
  names.arg=c("No consumption", "Consumption"))
active = table(cardio_train$gender)
barplot(active, main="Active Distribution",
  names.arg=c("Not active", "Active"))
cardio = table(cardio_train$cardio)
barplot(cardio, main="Cardio Distribution",
  names.arg=c("Presence of disease", "Absence of disease"))
```

We would not expect that the amount of people with cardiovascular diseases to be almost identical<br>
to the amount of those without, as such we can conclude that the data as almost certainly sampled to<br>
create an unbiased and equal distribution of this property.<br>
We would also expect that from random sampling the amount of men and women would be quite close, wich<br>
is not the case, with a lot more women than men.<br>
None of the features seem to follow a normal distribution wich may prove a problem. However due to the<br>
sample size such a problem is likely to be either mitigated or practically inexistent.<br>
The extremely skewed distriution of ap_hi, ap_lo, cholesterol and gluc might cause bias in the data.<br>


## Features normality test(qqnorm, test)

```{r featureNormality}
par(mfrow=c(2,2))
qqnorm(cardio_train$age,main="age"); qqline(cardio_train$age, col = 2,lwd=2,lty=2)
qqnorm(cardio_train$gender,main="gender"); qqline(cardio_train$gender, col = 2,lwd=2,lty=2)
qqnorm(cardio_train$height,main="height"); qqline(cardio_train$height, col = 2,lwd=2,lty=2)
qqnorm(cardio_train$weight,main="weight"); qqline(cardio_train$weight, col = 2,lwd=2,lty=2)
qqnorm(cardio_train$ap_hi,main="ap_hi"); qqline(cardio_train$ap_hi, col = 2,lwd=2,lty=2)
qqnorm(cardio_train$ap_lo,main="ap_lo"); qqline(cardio_train$ap_lo, col = 2,lwd=2,lty=2)
qqnorm(cardio_train$cholesterol,main="cholesterol"); qqline(cardio_train$cholesterol, col = 2,lwd=2,lty=2)
qqnorm(cardio_train$gluc,main="gluc"); qqline(cardio_train$gluc, col = 2,lwd=2,lty=2)
qqnorm(cardio_train$smoke,main="smoke"); qqline(cardio_train$smoke, col = 2,lwd=2,lty=2)
qqnorm(cardio_train$alco,main="alco"); qqline(cardio_train$alco, col = 2,lwd=2,lty=2)
qqnorm(cardio_train$active,main="active"); qqline(cardio_train$active, col = 2,lwd=2,lty=2)
qqnorm(cardio_train$cardio,main="cardio"); qqline(cardio_train$cardio, col = 2,lwd=2,lty=2)
library(tseries)
jarque.bera.test(cardio_train$age)
jarque.bera.test(cardio_train$gender)
jarque.bera.test(cardio_train$height)
jarque.bera.test(cardio_train$weight)
jarque.bera.test(cardio_train$ap_hi)
jarque.bera.test(cardio_train$ap_lo)
jarque.bera.test(cardio_train$cholesterol)
jarque.bera.test(cardio_train$gluc)
jarque.bera.test(cardio_train$smoke)
jarque.bera.test(cardio_train$alco)
jarque.bera.test(cardio_train$active)
jarque.bera.test(cardio_train$cardio)
```

From the plots and tests we can conclude that no feature is normaly distributed(or even close to it)<br>
wich will make the analisis more complicated.<br>

Other normality tests were tried, however all of them returned an error due to sample size, as such<br>
this one(the only one that worked) was used.<br>

## Feature correlations

```{r featureCorrelation}
cor(cardio_train)
```

We can observe that no two features are highly correlated with each other, with the correlation bettwen<br>
height and gender being the highest(0.5), wich on one hand will facilitate the analisis of the results, but<br>
on the othe, given that there is allways little correlation bettwen the features and target variable,<br>
obtaining good results may prove difficult.<br>


## Features scatterplots

```{r featureScatter}
pairs(cardio_train, pch = 19)
```

Due to the presence of a lot of categorical and binary variables most of the plots don't hold much meaning.<br>
There is however an interesting aspect of the data, in a lot of the plots, the points seem to get divided<br>
into two distinct groups. The meaning of this cannot be affered at this time, however this will be something<br>
to take into consideration as the work develops.<br>

## Response by value of categorical features

```{r histGender}
par(mfrow=c(2,1))
women =  cardio_train[which(cardio_train$gender==1),]$cardio
men =  cardio_train[which(cardio_train$gender==2),]$cardio
women = table(women)
barplot(women, main="Women Response",
  names.arg=c("Presence of disease", "Absence of disease"))
men = table(men)
barplot(men, main="Men Response",
  names.arg=c("Presence of disease", "Absence of disease"))
```

The results are almost exactly the same for men and women, as such this feature does not hold much information.<br>

```{r histCholesterol}
par(mfrow=c(3,1))
normalCholesterol =  cardio_train[which(cardio_train$cholesterol==1),]$cardio
highCholesterol =  cardio_train[which(cardio_train$cholesterol==2),]$cardio
veryHighColesterol =  cardio_train[which(cardio_train$cholesterol==3),]$cardio
normalCholesterol = table(normalCholesterol)
barplot(normalCholesterol, main="Normal cholesterol Response",
  names.arg=c("Presence of disease", "Absence of disease"))
highCholesterol = table(highCholesterol)
barplot(highCholesterol, main="High cholesterol Response",
  names.arg=c("Presence of disease", "Absence of disease"))
veryHighColesterol = table(veryHighColesterol)
barplot(veryHighColesterol, main="Very high cholesterol Response",
  names.arg=c("Presence of disease", "Absence of disease"))

```

The results are as we would theoritically expect, with higher amounts of cholesterol equating to higher chance<br>
of cardiovascular disease.<br>

```{r histGlucose}
par(mfrow=c(3,1))
normalGlucose =  cardio_train[which(cardio_train$gluc==1),]$cardio
highGlucose =  cardio_train[which(cardio_train$gluc==2),]$cardio
veryHighGlucose =  cardio_train[which(cardio_train$gluc==3),]$cardio
normalGlucose = table(normalGlucose)
barplot(normalGlucose, main="Normal glucose Response",
  names.arg=c("Presence of disease", "Absence of disease"))
highGlucose = table(highGlucose)
barplot(highGlucose, main="High glucose Response",
  names.arg=c("Presence of disease", "Absence of disease"))
veryHighGlucose = table(veryHighGlucose)
barplot(veryHighGlucose, main="Very high glucose Response",
  names.arg=c("Presence of disease", "Absence of disease"))
```

The results are as we would theoritically expect, with higher amounts of glucose equating to higher chance<br>
of cardiovascular disease.<br>

```{r histSmoke}
par(mfrow=c(2,1))
nonSmoker =  cardio_train[which(cardio_train$smoke==0),]$cardio
smoker =  cardio_train[which(cardio_train$smoke==1),]$cardio
nonSmoker = table(nonSmoker)
barplot(nonSmoker, main="Non smoker Response",
  names.arg=c("Presence of disease", "Absence of disease"))
smoker = table(smoker)
barplot(smoker, main="Smoker Response",
  names.arg=c("Presence of disease", "Absence of disease"))
```

The results are almost exactly the same for smokers and non smokers, as such this feature does not hold much<br>
information. This however goes against what is to be expected.<br>

```{r histAlcool}
par(mfrow=c(2,1))
noAlcoolIntake =  cardio_train[which(cardio_train$alco==0),]$cardio
alcoolIntake =  cardio_train[which(cardio_train$alco==1),]$cardio
noAlcoolIntake = table(noAlcoolIntake)
barplot(noAlcoolIntake, main="No alcool intake Response",
  names.arg=c("Presence of disease", "Absence of disease"))
alcoolIntake = table(alcoolIntake)
barplot(alcoolIntake, main="Alcool intake Response",
  names.arg=c("Presence of disease", "Absence of disease"))
```

The results are almost exactly the same for individuals who consume alcool as to those who do not, as such this<br>
feature does not hold much information. This however goes against what is to be expected.<br>

```{r histActive}
par(mfrow=c(2,1))
inactive =  cardio_train[which(cardio_train$active==0),]$cardio
active =  cardio_train[which(cardio_train$active==1),]$cardio
inactive = table(inactive)
barplot(inactive, main="Inactive Response",
  names.arg=c("Presence of disease", "Absence of disease"))
active = table(active)
barplot(active, main="Active Response",
  names.arg=c("Presence of disease", "Absence of disease"))
```

The results are almost exactly the same for for individuals that are active as well as those who are not as<br>
such this feature does not hold much information. This however goes against what is to be expected.<br>

## Simple processing

Some problems have been identified, namely the obviously wrong values in ap_hi and ap_lo, wich will now be<br>
taken care of.<br>

```{r outlierRemoval}
cardio_train_clean = cardio_train[which(!(cardio_train$ap_hi>500|cardio_train$ap_lo>500|cardio_train$ap_hi<0|cardio_train$ap_lo<0)),]
```

As other outliers have reasonable values they will be kept for now, and possibly removed later if they cause<br>
problems.<br>

We will also perform changes in some categorical variables.<br>
For gender we will change the values from 1-women, 2-men to a more standard 0-women, 1-men.<br>
For cholesterol we will turn the column type into factor.<br>
For gluc we will turn the column type into factor.<br>

```{r genderDecrease}
cardio_train_clean$gender = cardio_train_clean$gender-1
```

```{r cholesterolSplit}
cardio_train_clean$cholesterol = as.factor(cardio_train_clean$cholesterol)
```

```{r glucoseSplit}
cardio_train_clean$gluc = as.factor(cardio_train_clean$gluc)
```

## Null model

Now we will create the empty, null, model.<br>

```{r modelNull}
nullModel = lm(cardio_train_clean$cardio ~ 1)
nullModel
anova(nullModel)
```

From this model we obtained an intercept of 0.4949, wich means that the mean is 0.4949, wich in turn means<br>
that 49.49% of the individuals were found to have cardiovascular diseases.<br>
We also obtained a TSS of 17250 and a mean squared error of 0.24998

## Single model

Let's create models with a single variable to see their impact.<br>
For the case of the dummie vriables, these will also be analised separatly from each other.<br>

```{r modelSingle}
ageModel = lm(cardio_train_clean$cardio ~ cardio_train_clean$age)
ageModel
anova(ageModel)
anova(nullModel,ageModel)
genderModel = lm(cardio_train_clean$cardio ~ cardio_train_clean$gender)
genderModel
anova(genderModel)
anova(nullModel,genderModel)
heightModel = lm(cardio_train_clean$cardio ~ cardio_train_clean$height)
heightModel
anova(heightModel)
anova(nullModel,heightModel)
weightModel = lm(cardio_train_clean$cardio ~ cardio_train_clean$weight)
weightModel
anova(weightModel)
anova(nullModel,weightModel)
weightModel = lm(cardio_train_clean$cardio ~ cardio_train_clean$weight)
weightModel
anova(weightModel)
anova(nullModel,weightModel)
ap_hiModel = lm(cardio_train_clean$cardio ~ cardio_train_clean$ap_hi)
ap_hiModel
anova(ap_hiModel)
anova(nullModel,ap_hiModel)
ap_loModel = lm(cardio_train_clean$cardio ~ cardio_train_clean$ap_lo)
ap_loModel
anova(ap_loModel)
anova(nullModel,ap_loModel)
cholesterolModel = lm(cardio_train_clean$cardio ~ cardio_train_clean$cholesterol)
cholesterolModel
anova(cholesterolModel)
anova(nullModel,cholesterolModel)
glucModel = lm(cardio_train_clean$cardio ~ cardio_train_clean$gluc)
glucModel
anova(glucModel)
anova(nullModel,glucModel)
smokeModel = lm(cardio_train_clean$cardio ~ cardio_train_clean$smoke)
smokeModel
anova(smokeModel)
anova(nullModel,smokeModel)
alcoModel = lm(cardio_train_clean$cardio ~ cardio_train_clean$alco)
alcoModel
anova(alcoModel)
anova(nullModel,alcoModel)
activeModel = lm(cardio_train_clean$cardio ~ cardio_train_clean$active)
activeModel
anova(activeModel)
anova(nullModel,activeModel)
```

The minimum error obtained was that of 0.21 mean squared error from the regression with ap_hi.<br>

The gross effects obtained were:<br>
age: 4.855*10^-5<br>
gender: 0.007897<br>
height: -0.0006307<br>
weight: 0.006279<br>
ap_hi: 0.01105<br>
ap_lo: 0.01658<br>
cholesterol: 0.1609 for the second category and 0.3272 for the third category<br>
gluc: 0.1129 for the second category and 0.1425 for the third category<br>
smoke: -0.02883<br>
alco: -0.01802<br>
active: -0.04686<br>

## Full model

Now we will see the results with all of the variables.<br>

```{r modelFull}
fullModel = lm(cardio_train_clean$cardio ~ 
                 age +
                 gender +
                 height +
                 weight +
                 ap_hi +
                 ap_lo +
                 cholesterol +
                 gluc +
                 smoke +
                 alco +
                 active,
               data = cardio_train_clean
               )
fullModel
anova(fullModel)
anova(nullModel,fullModel)
```

With the saturated model we managed to decrease the error from the minimum of 0.21 to 0.20.<br>
There is however an obvious non significant feature.<br>
That being height.<br>
Given the nature of the problem, the intercept of -1.46 is a concerning value.<br>

## Model iteration

At this point we will iterate on the model until we reach the most statistically significant<br>

We will start from the saturated model, and remove one none significant variable.<br>
Then we will compute the new linear regression. And do the previous step untill all variables<br>
are significant.<br>
Then we will select from the model the most significant variable and add it. If it is significant<br>
go back to step one and repeat.<br>
After we only have significant variables and no significant one can be added we will check the cook's<br>
distances and leverages and remove observations with high values on both. If any are removed, go back<br>
to step one.<br>
After the variable selection and cook/leverage inspection we will check for confounding variables, by<br>
adding one variable at a time and checking the changes on the other ones. If a coeficient changes<br>
too much the variable will be added and we will go back to step one.<br>

## Iteration 1

First lets remove the variable with the highest p-value, the height feature.<br>

```{r modelIter1}
iterModel1 = lm(cardio_train_clean$cardio ~ 
                 age +
                 gender +
                 weight +
                 ap_hi +
                 ap_lo +
                 cholesterol +
                 gluc +
                 smoke +
                 alco +
                 active,
               data = cardio_train_clean
               )
iterModel1
anova(iterModel1)
anova(fullModel,iterModel1)
```

There are no more variables that are not significant, since we just removed the height, we wont<br>
try to put it back in.<br>

Since we now have all the significant variables, lets check the cook's distances.<br>

```{r modelIter2}
plot(hatvalues(iterModel1))
```

All the cook's distances are below 0.01, way below the required value to be concerning(1).<br>
Therefore it wont be required to look at the leverages.<br>

Now we will verify for confounding variables.<br>

```{r modelIter5}
iterModel2 = lm(cardio_train_clean$cardio ~ 
                 age +
                 gender +
                 height +
                 weight +
                 ap_hi +
                 ap_lo +
                 cholesterol +
                 gluc +
                 smoke +
                 alco +
                 active,
               data = cardio_train_clean
               )
iterModel2
anova(iterModel2)
anova(iterModel1,iterModel2)
```

Some of the values changed significantly and as such height is a confounding variable.<br>
Therefore it will be added back into the model.<br>

v)<br>

```{r ageCholesterolInteraction}
cardio_train_lowCol = cardio_train_clean[which(cardio_train_clean$cholesterol==1),]
cardio_train_medCol = cardio_train_clean[which(cardio_train_clean$cholesterol==2),]
cardio_train_highCol = cardio_train_clean[which(cardio_train_clean$cholesterol==3),]
lowColModel = lm(cardio_train_lowCol$cardio ~ 
                 age +
                 gender +
                 height +
                 weight +
                 ap_hi +
                 ap_lo +
                 gluc +
                 smoke +
                 alco +
                 active,
               data = cardio_train_lowCol
               )
lowColModel
medCol = lm(cardio_train_medCol$cardio ~ 
                 age +
                 gender +
                 height +
                 weight +
                 ap_hi +
                 ap_lo +
                 gluc +
                 smoke +
                 alco +
                 active,
               data = cardio_train_medCol
               )
medCol
highCol = lm(cardio_train_highCol$cardio ~ 
                 age +
                 gender +
                 height +
                 weight +
                 ap_hi +
                 ap_lo +
                 gluc +
                 smoke +
                 alco +
                 active,
               data = cardio_train_highCol
               )
highCol

calculate_mode <- function(x) {
  uniqx <- unique(na.omit(x))
  uniqx[which.max(tabulate(match(x, uniqx)))]
}

cardio_train_mean = cardio_train_lowCol
cardio_train_mean$gender = calculate_mode(cardio_train_lowCol$gender)
cardio_train_mean$height = mean(cardio_train_lowCol$height)
cardio_train_mean$weight = mean(cardio_train_lowCol$weight)
cardio_train_mean$ap_hi = mean(cardio_train_lowCol$ap_hi)
cardio_train_mean$ap_lo = mean(cardio_train_lowCol$ap_lo)
cardio_train_mean$smoke = calculate_mode(cardio_train_lowCol$smoke)
cardio_train_mean$alco = calculate_mode(cardio_train_lowCol$alco)
cardio_train_mean$active = calculate_mode(cardio_train_lowCol$active)
cardio_train_mean$cholesterol = calculate_mode(cardio_train_lowCol$cholesterol)
cardio_train_mean$gluc = calculate_mode(cardio_train_lowCol$gluc)

s = cardio_train_lowCol
cardio_train_lowCol = cardio_train_mean

plot(cardio_train_mean$age,cardio_train_mean$cardio)
pr1 <- predict(lowColModel, cardio_train_lowCol, interval="confidence")
yhat <- pr1[,1]
lines(cardio_train_lowCol$age, yhat, lwd=3, lty=1, col="magenta")

cardio_train_lowCol = s

cardio_train_mean = cardio_train_medCol
cardio_train_mean$gender = calculate_mode(cardio_train_medCol$gender)
cardio_train_mean$height = mean(cardio_train_medCol$height)
cardio_train_mean$weight = mean(cardio_train_medCol$weight)
cardio_train_mean$ap_hi = mean(cardio_train_medCol$ap_hi)
cardio_train_mean$ap_lo = mean(cardio_train_medCol$ap_lo)
cardio_train_mean$smoke = calculate_mode(cardio_train_medCol$smoke)
cardio_train_mean$alco = calculate_mode(cardio_train_medCol$alco)
cardio_train_mean$active = calculate_mode(cardio_train_medCol$active)
cardio_train_mean$cholesterol = calculate_mode(cardio_train_medCol$cholesterol)
cardio_train_mean$gluc = calculate_mode(cardio_train_medCol$gluc)

s = cardio_train_medCol
cardio_train_medCol = cardio_train_mean

pr1 <- predict(medCol, cardio_train_medCol, interval="confidence")
yhat <- pr1[,1]
lines(cardio_train_medCol$age, yhat, lwd=3, lty=1, col="blue")

cardio_train_medCol = s

cardio_train_mean = cardio_train_highCol
cardio_train_mean$gender = calculate_mode(cardio_train_highCol$gender)
cardio_train_mean$height = mean(cardio_train_highCol$height)
cardio_train_mean$weight = mean(cardio_train_highCol$weight)
cardio_train_mean$ap_hi = mean(cardio_train_highCol$ap_hi)
cardio_train_mean$ap_lo = mean(cardio_train_highCol$ap_lo)
cardio_train_mean$smoke = calculate_mode(cardio_train_highCol$smoke)
cardio_train_mean$alco = calculate_mode(cardio_train_highCol$alco)
cardio_train_mean$active = calculate_mode(cardio_train_highCol$active)
cardio_train_mean$cholesterol = calculate_mode(cardio_train_highCol$cholesterol)
cardio_train_mean$gluc = calculate_mode(cardio_train_highCol$gluc)

s = cardio_train_highCol
cardio_train_highCol = cardio_train_mean

pr1 <- predict(highCol, cardio_train_highCol, interval="confidence")
yhat <- pr1[,1]
lines(cardio_train_highCol$age, yhat, lwd=3, lty=1, col="red")

cardio_train_highCol = s
```

The previous plot indicated a significant change in slope(coefficient) of age for<br>
all three levels of cholesterol. Wich indicates significant interaction bettwen<br>
age and cholesterol.<br>

Let's now change our model to include that iteraction.<br>

```{r ageCholesterolModel}
ageColModel = lm(cardio_train_clean$cardio ~ 
                 age*
                 cholesterol  +
                 gender +
                 height +
                 weight +
                 ap_hi +
                 ap_lo +
                 gluc +
                 smoke +
                 alco +
                 active,
               data = cardio_train_clean
               )
ageColModel
anova(ageColModel)
anova(iterModel2,ageColModel)
```

The coefficients proved to be significant, proving our hypotesis.<br>

## Residuals

```{r modelResids}
plot(ageColModel$residuals)
boxplot(ageColModel$residuals,rstandard(ageColModel),rstudent(ageColModel))
qqnorm(ageColModel$residuals)
qqline(ageColModel$residuals)
hist(ageColModel$residuals)
```

From the plot the residuals seem to have homocedasticy, however from the normality plots<br>
we conclude that they are not normal, presenting a wave like pattern.<br>

## Further analisis of two variables

i)<br>
continuous variable: age<br>
categorical variable: cholesterol<br>

Gross effects:<br>
age - 0.00004855<br>
cholesterol2 - 0.1609<br>
cholesterol3 - 0.3272<br>
Adjusted efffects:<br>
age - 0.00003421<br>
cholesterol2 - 0.4711<br>
cholesterol3 - 0.6645<br>

The weight of the age decreased, wich means that some of the increase in the response explained by age<br>
is now being explained by other variables. This is corroborated by the fact that most other features have<br>
positive coefficients. This all means that an increase in age now has a greater increase in the value of the response.<br>
The weights of the cholesterol2 and cholesterol3 increased, wich means that some of the decrease in the<br>
response explained by them is now being explained by other variables. This is corroborated by the fact that<br>
most other features have positive coefficients. This all means that an increase in cholesterol now has a greater<br>
increase in the value of the response.<br>


ii)<br>
```{r agePredictions}
calculate_mode <- function(x) {
  uniqx <- unique(na.omit(x))
  uniqx[which.max(tabulate(match(x, uniqx)))]
}

cardio_train_mean = cardio_train_clean
cardio_train_mean$gender = calculate_mode(cardio_train_clean$gender)
cardio_train_mean$height = mean(cardio_train_clean$height)
cardio_train_mean$weight = mean(cardio_train_clean$weight)
cardio_train_mean$ap_hi = mean(cardio_train_clean$ap_hi)
cardio_train_mean$ap_lo = mean(cardio_train_clean$ap_lo)
cardio_train_mean$smoke = calculate_mode(cardio_train_clean$smoke)
cardio_train_mean$alco = calculate_mode(cardio_train_clean$alco)
cardio_train_mean$active = calculate_mode(cardio_train_clean$active)
cardio_train_mean$cholesterol = calculate_mode(cardio_train_clean$cholesterol)
cardio_train_mean$gluc = calculate_mode(cardio_train_clean$gluc)

s = cardio_train_clean
cardio_train_clean = cardio_train_mean

plot(cardio_train_clean$age,cardio_train_clean$cardio,ylim=c(-1,1.5))
pr1 <- predict(ageColModel, cardio_train_clean, interval="confidence")
pr2 <- predict(ageColModel, cardio_train_clean, interval="prediction")
yhat <- pr1[,1]
yhL1 <- pr1[,2]
yhU1 <- pr1[,3]
yhL2 <- pr2[,2]
yhU2 <- pr2[,3]
lines(cardio_train_clean$age, yhat, lwd=3, lty=1, col="magenta")
lines(cardio_train_clean$age, yhL1, lw=5, lt=3, col="blue")
lines(cardio_train_clean$age, yhU1, lw=5, lt=3, col="blue")
lines(cardio_train_clean$age, yhL2, lw=5, lt=3, col="green")
lines(cardio_train_clean$age, yhU2, lw=5, lt=3, col="green")

cardio_train_clean = s
```

We can observe an increase in the chance of cardiovascular disease as the age increases,<br>
as would be expected. However the prediction bands are outside the range of possible values<br>

iii)<br>
The coefficient for the cholesterol3 is 0.6645, while for cholesterol2 is 0.4711.<br>
Changing from cholesterol3 to cholesterol2 has a predicted change of -0.1888.<br>

```{r cholConfInts}
releveledCholesterol = relevel(cardio_train_clean$cholesterol,ref="3")
cholConfintModel = lm(cardio_train_clean$cardio ~ 
                 age *
                 releveledCholesterol  +
                 gender +
                 height +
                 weight +
                 ap_hi +
                 ap_lo +
                 gluc +
                 smoke +
                 alco +
                 active,
               data = cardio_train_clean
               )
cholConfintModel
confint(cholConfintModel, 'releveledCholesterol2', level=0.95)
confint(cholConfintModel, 'releveledCholesterol2', level=0.90)
```

The 95% interval for the change is [-0.3132855,-0.07350826].<br>
The 90% interval for the change is [-0.294010,-0.09278359].<br>
Meaning that there is about a 7-31% decrease in the chance of cardiovascular disease when<br>
going from the 3rd category to the 2nd one.<br>

iv)<br>

```{r ageEffects}
sdAge = sd(cardio_train_clean$age)
coefAge = ageColModel$coefficients[2]
change = sdAge*coefAge
change
```

An increase in age corresponding to a value of one of it's standard deviation causes an<br>
increase in the response of 0.08501982, meaning a 9% increase in the risk of cardiovascular<br>
disease.<br>